---
title: "Geo Boundary Data"
output: html_document
author: "Qing Chang"
date: "2025-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/PolBoundaryKG/")
library(DescTools)
library(geodata)
library(tidyverse)
library(states)
library(yaml)
library(countrycode)
library(r2country)
library(sf)
```

This document contains steps to construct spatial boundary knowledge files in CountryFiles folder. 

# Download Globael Data to Form a Country List

```{r}
world <- world(resolution=2, path = "~/PolBoundaryKG/R_Files/gadm")

world <- st_as_sf(world)

write_sf(world, "~/PolBoundaryKG/R_Files/gadm/world.shp")

code_list <- world %>% 
  st_drop_geometry() %>% 
  `$`(GID_0) %>%
  unique()

```


# Download Country Level Spatial Files Based on Country List 

```{r}
admin0 <- data.frame()

for (i in 1:length(code_list)) {
  
  print(code_list[i])
  
  tryCatch(
    {
      temp <- gadm(code_list[i], level=0, path = "~/Dropbox/PolBoundaryKG/R_Files/gadm")
      temp <- as.data.frame(temp)
      
      admin0 <- bind_rows(admin0,temp)} , error = function(e) {
        print("There is no admin0 in this country")
        
      } 
  )
  
  
}
```


GADM spatial boundaries include both UN recognized independent states and conflict territories. 
Therefore, we need to identify independent states using data from [REST](https://restcountries.com/).
We use variable "independent" in "countriesV3.1.json" stored in their [GitLab](https://gitlab.com/restcountries/restcountries/-/tree/master/src/main/resources?ref_type=heads). 

```{r}

json_data <- jsonlite::fromJSON("~/PolBoundaryKG/R_Files/countriesV3.1.json",
                                flatten=TRUE,
                                simplifyVector = F)

country_infors <- json_data %>% 
  map(unlist) %>% 
  map_df(bind_rows) %>% 
  select(GID_0 = cca3, is_independent = independent, continents) %>% 
  as.data.frame()

admin0 <- admin0 %>% 
  left_join(country_infors, 
            by = "GID_0")

admin0  <- admin0  %>% 
  mutate( 
    is_independent = if_else(COUNTRY == "Akrotiri and Dhekelia", "FALSE", is_independent), 
         is_independent = if_else(COUNTRY == "Caspian Sea", "FALSE", is_independent), 
         is_independent = if_else(COUNTRY == "Clipperton Island", "FALSE", is_independent), 
         is_independent = if_else(COUNTRY == "Kosovo", "FALSE", is_independent), 
         is_independent = if_else(COUNTRY == "Paracel Islands", "FALSE", is_independent),
         continents = if_else(COUNTRY == "Akrotiri and Dhekelia", "Europe", continents),
         continents = if_else(COUNTRY == "Caspian Sea", "Europe", continents), 
         continents = if_else(COUNTRY == "Clipperton Island", "Oceania", continents), 
         continents = if_else(COUNTRY == "Kosovo", "Europe", continents), 
         continents = if_else(COUNTRY == "Paracel Islands", "Asia", continents),
         #is_independent = if_else(str_detect(GID_0, "Z0"), "Disputed Area", is_independent),
    continents = if_else(is.na(continents) , "Asia", continents)
         ) %>% 
  filter(!str_detect(GID_0, "Z0"))


write_rds(admin0, "~/PolBoundaryKG/R_Files/gadm/admin0_gadm.rds")

```

# Download and Process Admin 1

```{r}

admin1 <- data.frame()

  
for (i in 1:length(code_list)) {
  
  print(code_list[i])
  
  tryCatch(
    {
    temp <- gadm(code_list[i], level=1, path = "~/Dropbox/PolBoundaryKG/R_Files/gadm")
    temp <- as.data.frame(temp)
    
    admin1 <- bind_rows(admin1,temp)} , error = function(e) {
      print("There is no admin1 in this country")
      
    } 
  )
  
  
}

```


We fixed naming and spatial boundary mistakes in Ukraine and British.


```{r}

### Ukraine first

UkR <- read_rds("~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_UKR_1_pk.rds")
UkR <- unwrap(UkR)
UkR <- st_as_sf(UkR)

missing_part <- UkR[which(UkR$GID_1=="?"), ]
join_part <- UkR[which(UkR$GID_1=="UKR.11_1"), ]
together <- st_union(join_part, missing_part)
together <- together[ ,-c(12:22)]
UkR <- UkR[ -which(UkR$GID_1=="?" | UkR$GID_1=="UKR.11_1"), ]

UkR <- bind_rows(UkR,together)
UkR <- vect(UkR)
UkR <- wrap(UkR)
write_rds(UkR, "~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_UKR_1_pk.rds")

### British second

GBR <- read_rds("~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_GBR_1_pk.rds")
GBR <- unwrap(GBR)
GBR <- st_as_sf(GBR)

GBR <- GBR %>% 
  mutate(NAME_1 = if_else(GID_1 == "GBR.1_1","England",NAME_1))

missing_part <- GBR[which(is.na(GBR$GID_1)), ]
join_part <- GBR[which(GBR$GID_1=="GBR.3_1"), ]
together <- st_union(join_part, missing_part)
together <- together[ ,-c(12:22)]
GBR <- GBR[ -which(is.na(GBR$GID_1) | GBR$GID_1=="GBR.3_1"), ]

GBR <- bind_rows(GBR,together)
GBR <- vect(GBR)
GBR <- wrap(GBR)
write_rds(GBR, "~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_GBR_1_pk.rds")


## Then correct values in admin1 files 
UKR <- read_rds("~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_UKR_1_pk.rds") 
UKR <- as.data.frame(unwrap(UKR))
GBR <- read_rds("~/PolBoundaryKG/R_Files/gadm/gadm/gadm41_GBR_1_pk.rds")
GBR <- as.data.frame(unwrap(GBR))

admin1 <- admin1 %>% 
  filter(GID_0!= "UKR" & GID_0!= "GBR") %>% 
  filter(!is.na(GID_0))

admin1 <- bind_rows(admin1, UKR,GBR)

write_rds(admin1, "~/PolBoundaryKG/R_Files/gadm/admin1_gadm.rds")

```


# Download and Process Admin 2


```{r}
admin2 <- data.frame()

for (i in 1:length(code_list)) {
  
  print(code_list[i])
  
  tryCatch(
    {
    temp <- gadm(code_list[i], level=2, path = "~/PolBoundaryKG/R_Files/gadm")
    temp <- as.data.frame(temp)
    
    admin2 <- bind_rows(admin2,temp)} , error = function(e) {
      print("There is no admin2 in country")
    
    } 
  )
  
  
}
```

We fixed naming and mistakes in Ukraine.

```{r}
## Correct Ukrainian  admin2 names

admin2 <- admin2 %>% 
  mutate(GID_1 = ifelse(GID_0 == "UKR"&GID_1=="?", "UKR.11_1", GID_1),
         NAME_1 = ifelse(GID_0 == "UKR" & NAME_1=="?", "Kiev City", NAME_1),
         NAME_2 = ifelse(GID_0 == "UKR" & NAME_2=="?", "Darnytskyi", NAME_2))

## Correct British admin2 names

admin2 <- admin2 %>% 
  mutate(GID_1 = ifelse(GID_2 == "GBR.1.6_1", "GBR.1_1", GID_1),
         GID_1 = ifelse(GID_2 == "GBR.3.27_1", "GBR.3_1", GID_1))



write_rds(admin2, "~/PolBoundaryKG/R_Files/gadm/admin2_gadm.rds")

```

# Create a File With All Admins and Spatial Geometry


```{r}
admin0_copy <- admin0 %>% 
  select(GID = GID_0, admin_name = COUNTRY) %>% 
  mutate(iso3c = GID,
         country_name = admin_name,
         admin_level = "Adm0",
         admin_new_name = paste0(GID, ".", "0", "_", admin_level)) 
  

admin1_copy <- admin1 %>% 
  select(GID = GID_1, admin_name = NAME_1, iso3c = GID_0, country_name=COUNTRY) %>% 
  mutate(
         admin_level = "Adm1",
         admin_new_name = paste0(GID, "_", admin_level)) 


admin2_copy <- admin2 %>% 
  select(GID = GID_2, admin_name = NAME_2, iso3c = GID_0, country_name=COUNTRY) %>% 
  mutate(
    admin_level = "Adm2",
    admin_new_name = paste0(GID, "_", admin_level)) 

admin_whole <- bind_rows(admin0_copy, admin1_copy, admin2_copy)

```


Now merge back with spatial polygon to add spatial geometry information.

```{r}
## Admin0 

filenames <- list.files("~/PolBoundaryKG/R_Files/gadm/gadm", pattern="0_pk.rds", full.names=TRUE)

admin0_shp <- lapply(filenames, function(i){
  temp <- read_rds(i)
  temp <- unwrap(temp)
  temp <- st_as_sf(temp)
  return(temp)
}) %>% bind_rows()

admin0_shp <- admin0_shp %>% 
  select(GID = GID_0)

admin0_shp <- merge(admin0_shp, admin0_copy, by = "GID")

admin0_shp <- admin0_shp %>% 
  select(-admin_new_name)

write_rds(admin0_shp, "~/PolBoundaryKG/R_Files/gadm/admin_whole_shp/admin0_shp.rds")

## Admin1 

filenames <- list.files("~/PolBoundaryKG/R_Files/gadm/gadm", pattern="1_pk.rds", full.names=TRUE)

admin1_shp <- lapply(filenames, function(i){
  temp <- read_rds(i)
  temp <- unwrap(temp)
  temp <- st_as_sf(temp)
  return(temp)
}) %>% bind_rows()

admin1_shp <- admin1_shp %>% 
  select(GID = GID_1)

admin1_shp <- merge(admin1_shp, admin1_copy, by = "GID")

admin1_shp <- admin1_shp %>% 
  select(-admin_new_name)


write_rds(admin1_shp, "~/PolBoundaryKG/R_Files/gadm/admin_whole_shp/admin1_shp.rds")

## Admin2

filenames <- list.files("~/PolBoundaryKG/R_Files/gadm/gadm", pattern="_2_pk.rds", full.names=TRUE)

admin2_shp <- lapply(filenames, function(i){
  temp <- read_rds(i)
  temp <- unwrap(temp)
  temp <- st_as_sf(temp)
  return(temp)
}) %>% bind_rows()

admin2_shp <- admin2_shp %>% 
  select(GID = GID_2)

admin2_shp <- merge(admin2_shp, admin2_copy, by = "GID")

admin2_shp <- admin2_shp %>% 
  select(-admin_new_name)


write_rds(admin2_shp, "~/PolBoundaryKG/R_Files/gadm/admin_whole_shp/admin2_shp.rds")
```

There are many missing administration names across different administration levels, we use 
[GeoNames Web Services](https://www.geonames.org/export/ws-overview.html) to find missing.
In order to use this service, you need to first register [here](https://www.geonames.org/login), and 
remember your username to be used for each request.


```{r}
################ Step1: 
############### get a list of data missing admin names and save for later

admin_whole_shp <- bind_rows(admin0_shp, admin1_shp, admin2_shp)

admin_missing <- admin_whole_shp %>% 
  filter(is.na(admin_name) | str_detect(admin_name,"[\\?]"))


################ Step2: 
############### read functions to send request and receive data 

source("GNfindNearby.R")

################ Step3: 
############### Using geonames web services to find missing
options(geonamesUsername= "\Your_User_Name")
sf::sf_use_s2(FALSE)

coord <- missing_names %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  as.data.frame()

coord$level <- ifelse(missing_names$admn_lv == "Adm1", "ADM1", "ADM2")
coord$GID <- missing_names$GID
coord$web_name <- NA

for (i in 1:nrow(coord)) {
  
  tryCatch(
    {
      temp <- suppressWarnings(GNfindNearby(lat = coord$Y[i], lng = coord$X[i], featureCode = coord$level[i]))
      
      coord$web_name[i] <- temp[["geonames"]][[1]][["name"]]
      
      print(paste0(i, " Finish!" ))
      
    } , error = function(e) {
      print(paste0("There is a mistake in ", i ))
      
    } 
  )
  
}

missing_names <- missing_names %>% 
  st_drop_geometry() %>% 
  select(GID, iso3c, country_name = cntry_n, admin_level = admn_lv) %>% 
  mutate( admin_new_name = paste0(GID, "_", admin_level))


coord <- coord %>% 
  select(GID, admin_name = web_name) %>% 
  left_join(missing_names, by = "GID")


# Remove missing names and add back geonames ones 

admin_whole <- admin_whole %>% 
  filter(!is.na(admin_name) ) %>% 
  filter(!str_detect(admin_name,"[\\?]"))

admin_whole <- bind_rows(admin_whole, coord)
```

# Create Concept and Edge Files

```{r}
# Create file for concept

admin_whole <- admin_whole %>% 
  mutate(admin_id = str_replace_all(country_name, fixed(" "), ""),
       admin_id = str_replace_all(admin_id, "[\r\n]" , ""),
       admin_id = str_replace_all(admin_id, "," , "")
       )


# Create file for relation

admin01_relation <- admin1 %>% 
  select(GID_0, GID_1, country_name = COUNTRY) %>% 
  mutate(
    iso3c = GID_0,
    GID_0 = paste0(GID_0, ".", "0", "_", "Adm0"),
    GID_1 = paste0(GID_1, "_", "Adm1")) %>% 
  mutate(relation = paste0("(",GID_0, ", ", GID_1, ")")) %>% 
  mutate(admin_id = str_replace_all(country_name, fixed(" "), ""),
         admin_id = str_replace_all(admin_id, "[\r\n]" , ""),
         admin_id = str_replace_all(admin_id, "," , "")
  ) %>% 
  select(iso3c, relation,  admin_id)


admin12_relation <- admin2 %>% 
  select(GID_1, GID_2, iso3c = GID_0, country_name = COUNTRY) %>% 
  mutate(
    GID_1 = paste0(GID_1, "_", "Adm1"),
    GID_2 = paste0(GID_2, "_", "Adm2")) %>% 
  mutate(relation = paste0("(",GID_1, ", ", GID_2, ")")) %>% 
  mutate(admin_id = str_replace_all(country_name, fixed(" "), ""),
         admin_id = str_replace_all(admin_id, "[\r\n]" , ""),
         admin_id = str_replace_all(admin_id, "," , "")
  ) %>% 
  select(iso3c, relation,  admin_id)


admin_relation_all <- bind_rows(admin01_relation,admin12_relation)

## begin to create country files

path <- "~/PolBoundaryKG/CountryFiles/"

name_list <- unique(admin_relation_all$admin_id)


for (i in 1:length(name_list)) {
  
  
  name_concept <- paste0(path, name_list[i], "Concepts.yaml")
  name_relation <- paste0(path, name_list[i], "Edges.yaml")
  
  df <- admin_whole %>% 
    filter(admin_id == name_list[i]) %>% 
    select(-admin_id)
  
  
  df_concept <- list(concepts=split(replace(df, "admin_new_name", NULL), df$admin_new_name))
  
  write_yaml(df_concept, name_concept)
  
  
  df_relation <- admin_relation_all %>% 
    filter(admin_id == name_list[i])
  
  
  write_yaml(data.frame(relations= df_relation$relation), indent = 4, 
             indent.mapping.sequence =T, name_relation)
  
}

## create global concept

global_data <-  admin0 %>% 
  select(GID = GID_0, country_name = COUNTRY, continents) %>% 
  mutate(iso3c = GID,
         admin_new_name = paste0(GID, ".", "0", "_", "Adm0")) %>% 
  mutate(continent = str_replace_all(continents, fixed(" "), ""),
         continent = str_replace_all(continent, "[\r\n]" , ""),
         continent = str_replace_all(continent, "," , "")
  ) %>% 
  mutate( continent = paste0(continent, "Continent")) 

## step 1: add country nodes 
country_node=split(replace(global_data %>% select(-continent), "admin_new_name", NULL), global_data$admin_new_name)

## step 2: add continents nodes 
conten <- global_data %>% 
  group_by(continent, continents) %>% 
  distinct(continent) %>% 
  ungroup() %>% 
  rename(Continent_name = continents)

conten_node = split(replace(conten, "continent", NULL), conten$continent)

df_concept <- list(concepts = append(conten_node, country_node))

write_yaml(df_concept, "~/PolBoundaryKG/GlobalConcepts.yaml")


## create global relation


global_data <-  global_data %>% 
  mutate(relation = paste0("(",continent, ", ",admin_new_name, ")")) 

write_yaml(data.frame(relations= global_data $relation), indent = 4, 
           indent.mapping.sequence =T, "~/PolBoundaryKG/GlobalEdges.yaml")
```


